Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-10-16 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: minor edits to cover letter and SI Appendix

| Phenology type | Subpopulation | Effects Selected per GWAS, Strong Subset | Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
|:-----------|-----------:|-----------:|-----------:|------------|------------|
| Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
|                |       Midwest |                                    33000 |                                    66000 | 57729                        | 115458                       |
|                |          Both |                                    19000 |                                    38000 | 44765                        | 89530                        |
| Flowering Date |          Gulf |                                    19000 |                                    38000 | 50490                        | 100980                       |
|                |       Midwest |                                    33000 |                                    66000 | 63367                        | 126734                       |
|                |          Both |                                    19000 |                                    38000 | 35735                        | 71470                        |

@@

| Greedy Algorithm Iteration \# |     Gulf Greenup Date |     Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
|:----------|----------:|----------:|----------:|-----------|-----------|-----------|
| 1                             |              Identity |                 Identity |             Identity | Identity                | Identity                       | Identity                |
| 2                             |         Simple Het. 2 |         Single Effect OK |     Single Effect MI | Simple Het. 1           | Simple Het. 1                  | Single Effect TX3       |
| 3                             |      Single Effect MO |            Equal Effects |     Single Effect SD | Single Effect MI        | Single Effect MI               | Single Effect TX1       |
| 4                             |      Single Effect NE |        Single Effect TX3 |     Single Effect NE | Single Effect TX3       | Single Effect TX2              | Single Effect OK        |
| 5                             | Daylength (14d prior) |        Single Effect TX2 |    Single Effect TX2 | Single Effect SD        | Single Effect MO               | Single Effect MI        |
| 6                             |      Single Effect MI |  Cumulative GDD, 12C, 1d |     Single Effect OK | Single Effect OK        | Single Effect TX3              | Single Effect NE        |
| 7                             |      Single Effect SD | Cumulative GDD, 12C, 28d |    Single Effect TX3 | Single Effect NE        | Daylength change (s), 2d prior | Single Effect TX2       |
| 8                             |     Single Effect TX2 |                       \- |        Simple Het. 1 | Single Effect TX1       | Single Effect SD               | Single Effect MO        |
| 9                             |    Cumulative GDD, 6d |                       \- |    Single Effect TX1 | Single Effect TX2       | \-                             | Single Effect SD        |
| 10                            |     Single Effect TX3 |                       \- |     Single Effect MO | Single Effect MO        | \-                             | Simple Het. 1           |
| 11                            |   Data-driven $PCA_2$ |                       \- |         Temp Ave, 1d | Cumulative rainfall, 7d | \-                             | Cumulative rainfall, 1d |
| 12                            |                    \- |                       \- | Daylength change (s) | \-                      | \-                             | \-                      |

>>>>

| Phenology type | Subpopulation | Effects Selected per GWAS, Strong Subset | Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
|:---------------|--------------:|-----------------------------------------:|-----------------------------------------:|------------------------------|------------------------------|
| Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
|                |       Midwest |                                    33000 |                                    66000 | 57729                        | 115458                       |
|                |          Both |                                    19000 |                                    38000 | 44765                        | 89530                        |
| Flowering Date |          Gulf |                                    19000 |                                    38000 | 50490                        | 100980                       |
|                |       Midwest |                                    33000 |                                    66000 | 63367                        | 126734                       |
|                |          Both |                                    19000 |                                    38000 | 35735                        | 71470                        |

@@

| Greedy Algorithm Iteration \# |     Gulf Greenup Date |     Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
|:------------------------------|----------------------:|-------------------------:|---------------------:|-------------------------|--------------------------------|-------------------------|
| 1                             |              Identity |                 Identity |             Identity | Identity                | Identity                       | Identity                |
| 2                             |         Simple Het. 2 |         Single Effect OK |     Single Effect MI | Simple Het. 1           | Simple Het. 1                  | Single Effect TX3       |
| 3                             |      Single Effect MO |            Equal Effects |     Single Effect SD | Single Effect MI        | Single Effect MI               | Single Effect TX1       |
| 4                             |      Single Effect NE |        Single Effect TX3 |     Single Effect NE | Single Effect TX3       | Single Effect TX2              | Single Effect OK        |
| 5                             | Daylength (14d prior) |        Single Effect TX2 |    Single Effect TX2 | Single Effect SD        | Single Effect MO               | Single Effect MI        |
| 6                             |      Single Effect MI |  Cumulative GDD, 12C, 1d |     Single Effect OK | Single Effect OK        | Single Effect TX3              | Single Effect NE        |
| 7                             |      Single Effect SD | Cumulative GDD, 12C, 28d |    Single Effect TX3 | Single Effect NE        | Daylength change (s), 2d prior | Single Effect TX2       |
| 8                             |     Single Effect TX2 |                       \- |        Simple Het. 1 | Single Effect TX1       | Single Effect SD               | Single Effect MO        |
| 9                             |    Cumulative GDD, 6d |                       \- |    Single Effect TX1 | Single Effect TX2       | \-                             | Single Effect SD        |
| 10                            |     Single Effect TX3 |                       \- |     Single Effect MO | Single Effect MO        | \-                             | Simple Het. 1           |
| 11                            |   Data-driven $PCA_2$ |                       \- |         Temp Ave, 1d | Cumulative rainfall, 7d | \-                             | Cumulative rainfall, 1d |
| 12                            |                    \- |                       \- | Daylength change (s) | \-                      | \-                             | \-                      |

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.qmd
revision2: minor edits to cover letter and SI Appendix

Our work is a key methodological development in understanding genetic variation in genotype-by-environment interactions, or GxE. GxE interactions must vary across the genome; that is, different genes and genetic variants can exhibit different responses to the environment. Yet current statistical models to test for GxE don't have robust frameworks to specify and compete different models of GxE effects, instead focusing on finding a single, best fitting form of GxE covariance. We use multivariate adaptive shrinkage (mash [@urbut2018]) to determine the regions of the switchgrass genome with different types of GxE, specifically regions of the genome with effects on phenology that covary with several weather-based 'cues'. We use a greedy algorithm to compete and select types of GxE and GxWeather that improve our models, then use mash to identify which SNP effects are best modeled with each type of GxE or GxWeather.

>>>>

Our work is a key methodological development in understanding genetic variation in genotype-by-environment interactions, or GxE. GxE interactions must vary across the genome; that is, different genes and genetic variants can exhibit different responses to the environment. Yet current statistical models to test for GxE don't have robust frameworks to specify and compete different models of GxE effects, instead focusing on finding a single, best fitting form of GxE covariance. We use multivariate adaptive shrinkage (mash, [Urbut *et al*., 2018](https://doi.org/10.1038/s41588-018-0268-8)) to determine the regions of the switchgrass genome with different types of GxE, specifically regions of the genome with effects on phenology that covary with several weather-based 'cues'. We use a greedy algorithm to compete and select types of GxE and GxWeather that improve our models, then use mash to identify which SNP effects are best modeled with each type of GxE or GxWeather.

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.tex
revision2: minor edits to cover letter and SI Appendix

Our work is a key methodological development in understanding genetic
variation in genotype-by-environment interactions, or GxE. GxE
interactions must vary across the genome; that is, different genes and
genetic variants can exhibit different responses to the environment. Yet
current statistical models to test for GxE don't have robust frameworks
to specify and compete different models of GxE effects, instead focusing
on finding a single, best fitting form of GxE covariance. We use
multivariate adaptive shrinkage (mash {[}@urbut2018{]}) to determine the
regions of the switchgrass genome with different types of GxE,
specifically regions of the genome with effects on phenology that covary
with several weather-based `cues'. We use a greedy algorithm to compete
and select types of GxE and GxWeather that improve our models, then use
mash to identify which SNP effects are best modeled with each type of
GxE or GxWeather.

>>>>

Our work is a key methodological development in understanding genetic
variation in genotype-by-environment interactions, or GxE. GxE
interactions must vary across the genome; that is, different genes and
genetic variants can exhibit different responses to the environment. Yet
current statistical models to test for GxE don't have robust frameworks
to specify and compete different models of GxE effects, instead focusing
on finding a single, best fitting form of GxE covariance. We use
multivariate adaptive shrinkage (mash,
\href{https://doi.org/10.1038/s41588-018-0268-8}{Urbut \emph{et al}.,
2018}) to determine the regions of the switchgrass genome with different
types of GxE, specifically regions of the genome with effects on
phenology that covary with several weather-based `cues'. We use a greedy
algorithm to compete and select types of GxE and GxWeather that improve
our models, then use mash to identify which SNP effects are best modeled
with each type of GxE or GxWeather.
